### 2022-11-09 NSF2022/Description/specificaim2.tex
new refs

%\label{subsec:specific_aim_2}
%\begin{wrapfigure}[14]{r}{0.61\textwidth}
%  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
%    \vspace{-15pt}
%    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
%    amphiphilic particles suspended in a solvent. The solvent velocity
%    satisfies the Stokes equations, and the hydrophobic forces depend on
%    the solvent activity that satisfies the screened Laplace equation.
%    The geometry is updated by solving the mobility problem.} 
%\end{wrapfigure}
The HAP model requires solving exterior problems for the order
parameter~\eqref{eqn:phase} and the Stokes equations~\eqref{eqn:stokes}.
We will develop methods based on combinations of volume and layer
potentials, and inexact Newton's method. As mentioned in
Section~\ref{sec:bie}, numerical methods to solve these integral
equations (IEs) have several advantages over their PDE-based
counterpart. The PIs have demonstrated that IEs are a powerful tool to
simulate two-dimensional suspensions of amphiphilic
particles~\cite{Fu2018_SIAM, FuQuRyYo22, fu-ryh-qua-you2022}. Moreover,
elements of these formulations have been described in three
dimensions~\cite{ying_2006, manasthesis, rac-gre2016}. This proposal
will extend the two-dimensional methods to more general amphiphilic
particle models and also to three dimensions. In addition, the methods
will allow us to use results we have shown for $\mathcal{O}(100)$
granules, and extend them to suspensions with 100X--1000X more granules.
These methods will be motivated by models and analysis developed in
Specific Aims 1 and 3. \S\ref{subsec:AC} describes how the non-linear
equation for the order parameter~\eqref{eqn:phase} will be iteratively
solved using an IE formulation. Then, \S\ref{subsec:NumericalIssues}
describes how we will combine accurate quadrature and fast summation
methods to accelerate the time to accurate solutions. A natural
extension to create simulations that are realizable by experimentalists
is to include confinement, and this is discussed in
\S\ref{subsec:Confinement}. This extension will also allow us to study
new rheological properties of the suspensions.  Finally,
\S\ref{subsec:timeStepping} proposes collision-aware adaptive time
stepping methods.

@@

The PIs have developed numerical methods to solve~\eqref{eqn:phase} when
$f(u) = u^2$---this choice results in the linear screened Laplace
equation. For a more general double well potential $f$, with local
minimums at $u_0$ and $u_1$, we will interpret the solution
of~\eqref{eqn:phase} as the root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  J[u] = -\rho^2 \Delta u + \tfrac{1}{2}f'(u).
\end{align}
This non-linear PDE will be solved with Newton's method, $u^{N+1} =
u^{N} - \alpha (dJ^N)^{-1} J(u^N)$, where $dJ^N = -\rho^2 \Delta +
\tfrac{1}{2}f''(u^N)$ is the Fr\'{e}chet derivative of $F$, and $\alpha$
is the step size which will be determined using backtracking. As
expected, the biggest computational challenge is solving $-\rho^2 \Delta
v + \tfrac{1}{2}f''(u^{N}) v = J[u^N]$. Solving this variable
coefficient PDE with an integral equation is challenging since a
fundamental solution is not readily available. However, by choosing a
double well potential with $f''(u_0) = f''(u_1) = \beta$, we propose to
instead use an inexact Newton's method where $dJ^{-1}$ is approximated
by solving $-\rho^2 \Delta v + \tfrac{beta}{2} v = J[u^N]$. We test the
efficacy of exact and inexact Newton's method applied
to~\eqref{eqn:phaseRoot} in $(0,1)$ by discretizing $dJ$ with a
second-order finite difference method. The double well potential is
illustrated in Figure~\ref{fig:CA}(a), and both the exact and inexact
Newton's methods converge to the solution in Figure~\ref{fig:CA}(b).
Figure~\ref{fig:CA}(c) shows the max norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach of a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.
\todo[inline]{put pics here}

%\begin{wrapfigure}[27]{l}{1.8in}
%  \vspace{-4pt}
%  \centering
%  \input{figures/doublewell.tikz} 
%  \caption{\label{fig:CA} (a) The double well potential $u_0 =
%  0.5$, $u_1 = 2$, $a = 1$, $b = 0.4$. (b) The solution of the
%  Allen-Cahn equation~\eqref{eq:SL}. (c) The convergence of the exact
%  and inexact Newton's methods.}
%\end{wrapfigure}
Having demonstrated that the inexact Newton method converges in one
dimension, it will be coupled with IE formulations to 
solve~\eqref{eq:SL} in two and three dimensions. We first partition
$\Omega$ into $\Omega_0 = \{\xx \in \Omega \:|\: u(\xx) < (u_0 +
u_1)/2\}$ and $\Omega_1 = \{\xx \in \Omega \:|\: u(\xx) > (u_0 +
u_1)/2\}$, and we let $\Lambda = \{\xx \in \Omega \:|\: u(\xx) = (u_0 +
u_1)/2 \}$ be their common boundary, and we let $G_0$ be the fundamental
solution of $(-\rho^2 \Delta + a)$ and $G_1$ be the fundamental
solution of $(-\rho^2 \Delta + 1)$. Then, the solution
of~\eqref{eq:screenedPoisson} can be written as $u = u^P + u^H$ where
$u^P$ is a volume potential and $u^H$ is a layer potential. The volume
potential is
\begin{align*}
  u^{P}(\xx) = \int_{\Omega_i} G_i(\xx,\yy) J[u^N](\yy)\, d\yy, 
    \quad i = 1,2,
\end{align*}
but it is discontinuous across $\Lambda$ and does not satisfy the
prescribed hydrophobic boundary conditions on the particles $\Gamma_i$.
Assuming the particles are entirely contained in the hydrophobic region,
the boundary conditions will be enforced with the layer potentials
\begin{align*}
  u^{H}(\xx) = \begin{cases}
    \displaystyle\int_{\Lambda} G_0(\xx,\yy) \eta_0(\yy) \, ds_\yy + 
    \displaystyle\int_{\Gamma} \pderiv{}{\nn_\yy} G_0(\xx,\yy)
      \sigma(\yy) \,ds_\yy, \quad & \xx \in \Omega_0, \\
    \displaystyle\int_{\Lambda} G_1(\xx,\yy) \eta_1(\yy) \, ds_\yy,
     & \xx \in \Omega_1.
  \end{cases}
\end{align*}
The density functions $\sigma$, $\eta_0$, and $\eta_1$ are determined so
that $u = u^P + u^H$ satisfies the hydrophobic boundary condition on
$\Gamma = \cup_i \Gamma_i$, and that it is continuous and differentiable
across $\Lambda$. This approach of using IEs to solve elliptic PDEs with
piecewise constant coefficients has successfully been used in other
application such as scattering~\cite{hyu-bar2014, che-cho-cai2018}.

>>>>

%\label{subsec:specific_aim_2}
%\begin{wrapfigure}[14]{r}{0.61\textwidth}
%  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
%    \vspace{-15pt}
%    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
%    amphiphilic particles suspended in a solvent. The solvent velocity
%    satisfies the Stokes equations, and the hydrophobic forces depend on
%    the solvent activity that satisfies the screened Laplace equation.
%    The geometry is updated by solving the mobility problem.} 
%\end{wrapfigure}
The HAP model requires solving exterior problems for the order
parameter~\eqref{eqn:phase} and the Stokes equations~\eqref{eqn:stokes}.
We will develop methods based on combinations of volume and layer
potentials, and inexact Newton's method. As mentioned in
Section~\ref{sec:bie}, numerical methods to solve these integral
equations (IEs) have several advantages over their PDE-based
counterpart. The PIs have demonstrated that IEs are a powerful tool to
simulate two-dimensional suspensions of amphiphilic
particles~\cite{Fu2018_SIAM, FuQuRyYo22, fu-ryh-qua-you2022}, and
elements of these formulations have been developed in three
dimensions~\cite{ying_2006, manasthesis, rac-gre2016}. This proposal
will extend the two-dimensional methods to more general amphiphilic
particle models and also to three dimensions. In addition, the methods
will allow us to use results we have shown for $\mathcal{O}(100)$
granules, and extend them to suspensions with 100X--1000X more granules.
These methods will be motivated by models and analysis developed in
Specific Aims 1 and 3. \S\ref{subsec:AC} describes how the non-linear
equation for the order parameter~\eqref{eqn:phase} will be iteratively
solved using an IE formulation. Then, \S\ref{subsec:NumericalIssues}
describes how we will combine accurate quadrature and fast summation
methods to accelerate the time to accurate solutions. A natural
extension to create simulations that are realizable by experimentalists
is to include confinement, and this is discussed in
\S\ref{subsec:Confinement}. This extension will also allow us to develop
new rheological properties of the suspensions.  Finally,
\S\ref{subsec:timeStepping} proposes collision-aware adaptive time
stepping methods.

@@

\begin{wrapfigure}[28]{l}{1.8in}
  \vspace{-8pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} \footnotesize (a) A double well potential. (b)
  The solution of the one-dimensional phase
  transition~\eqref{eqn:phase}. (c) The convergence of exact and inexact
  Newton's methods.}
\end{wrapfigure}
The PIs have developed numerical methods to solve~\eqref{eqn:phase} when
$W(\phi) = \phi^2$---this choice results in the linear screened Laplace
equation. For a general double well potential $W$, with local minima at
$\phi_0$ and $\phi_1$, we will interpret the solution
of~\eqref{eqn:phase} as the root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  J[\phi] = -\epsilon^2 \Delta \phi + \tfrac{1}{2}W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \alpha (dJ^N)^{-1} J(\phi^N)$, where $dJ^N =
-\epsilon^2 \Delta + \tfrac{1}{2}W''(\phi^N)$ is the Fr\'{e}chet
derivative of $F$, and $\alpha$ is the step size that will be determined
using backtracking. As expected, the biggest computational challenge is
solving $-\epsilon^2 \Delta v + \tfrac{1}{2}W''(\phi^{N}) v =
J[\phi^N]$. Solving this variable coefficient PDE with an integral
equation is challenging since a fundamental solution is not readily
available. However, by choosing a double well potential with
$W''(\phi_0) = W''(\phi_1) = \beta$, we propose to instead use an
inexact Newton's method where $dJ^{-1}$ is approximated by solving
$-\epsilon^2 \Delta v + \tfrac{\beta}{2} v = J[\phi^N]$. We test the
efficacy of exact and inexact Newton's method applied
to~\eqref{eqn:phaseRoot} in $(0,1)$ by discretizing $dJ$ with a
second-order finite difference method. The double well potential is
illustrated in Figure~\ref{fig:CA}(a), and both the exact and inexact
Newton's methods converge to the solution in Figure~\ref{fig:CA}(b).
Figure~\ref{fig:CA}(c) shows the max norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach of a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.

Having demonstrated that the inexact Newton method converges in one
dimension, it will be coupled with IE formulations to
solve~\eqref{eqn:phase} in two and three dimensions. The solution of
$-\epsilon^2 \Delta v + \tfrac{\beta}{2} v = J[\phi^N]$ can be written
as $v = \phi^P + \phi^H$ where $\phi^P$ is a volume potential and
$\phi^H$ is a layer potential. The volume potential is
\begin{align*}
  \phi^{P}(\xx) = \int_{\Omega(t)} G(\xx,\yy) J[\phi^N](\yy)\, d\yy, 
    \quad \xx \in \Omega(t),
\end{align*}
where $G$ is the fundamental solution of $-\epsilon^2 \Delta +
\tfrac{\beta}{2}$. Volume potentials are more challenging to compute
numerically than layer potentials. However, very recently, a suite of
new methods have been developed to accurately and efficiently compute
volume potentials~\cite{fry-leh-tor2018, and-zhu-vee2022}. The boundary
condition will be enforced using $\phi^H(\xx)$ which will be represented
as a double-layer potential. This approach of combining volume and layer
potentials has successfully been used by PI BQ~\cite{kro-qua2010} and
others~\cite{fry-kro-tor2019, kli-ask-kro2020, che-hua-lei2005}.
